---
title: "Factors, Strings, and Dates"
output:
  prettydoc::html_pretty:
    theme: tactile
    highlight: github
    css: styles.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options("couRsework-img_path" = here::here("Images", "icons")) 
### change this if icons local
```

```{r, include = FALSE}
library(tidyverse)
library(couRsework)
library(flair)
library(lubridate)
```


```{css, echo = FALSE}
img {
  vertical-align:middle;
}

body {
  text-shadow:none;
}



```{r, results = "asis", echo = FALSE}
make_summary_table(vid_time = "30 min", 
                   reading_time = "0-30 min", 
                   work_time = "120 min", 
                   num_activities = 6)
```

---

# Factors

```{r, results = "asis", echo = FALSE}
req_vid("Factors with forcats", type = "Required", url = "https://www.youtube.com/embed/bBfZFasd61c")
```

```{r, results = "asis", echo = FALSE}
req_read("R4DS Chapter 15: Factors", type = "Recommended", url = "https://r4ds.had.co.nz/factors.html")
```

---

```{r, results = "asis", echo = FALSE}
checkin("Functions from forcats")
```

Answer the questions from lecture:

* What is the difference between `fct_relevel()` and `fct_recode()`?

* What is the difference between `fct_collapse()` and `fct_lump()`?

* What is the difference between `fct_reorder()` and `fct_inorder()`?

*Feel free to share answers with your classmates in the Discord server, to make sure you understand.*

```{r, results = "asis", echo = FALSE}
checkin("Apply your skills")
```

In your *colleges_clean* dataset, adjust the `REGION` variable to have nicer or clearer labels and/or better categories.

---

# Strings

```{r, results = "asis", echo = FALSE}
req_vid("Strings with stringr", type = "Required", url = "https://www.youtube.com/embed/Oo_MtXwpi2Q")
```

```{r, results = "asis", echo = FALSE}
req_read("R4DS Chapter 14: Strings", type = "Recommended", url = "https://r4ds.had.co.nz/strings.html")
```

---

```{r, results = "asis", echo = FALSE}
checkin("Functions from stringr")
```

Answer the questions from lecture:

* What is the difference between `length()` and `str_length()`?

* What is the difference between `str_replace()` and `str_replace_all()`?

* What is the difference between `str_trim()` and `str_trunc()`?

* What is the difference between `str_sub()` and `str_subset()`?

* What does the `collapse` argument do in `str_c()`?

*Feel free to share answers with your classmates in the Discord server, to make sure you understand.*

```{r, results = "asis", echo = FALSE}
checkin("Apply your skills")
```

Using your *colleges_clean* dataset, answer these questions:

**Question 1:** How *exactly* does Cal Poly's name appear in this dataset?

**Question 2:** How many colleges have the word "University" in their name?

**Question 3:** Which college has the *longest* name?  The *shortest* name?

---

# Dates

```{r, results = "asis", echo = FALSE}
req_vid("DateTime Objects with lubridate", type = "Required", url = "https://www.youtube.com/embed/lmDL4o1aa0Q")
```


```{r, results = "asis", echo = FALSE}
req_read("R4DS Chapter 16: Dates and Times", type = "Recommended", url = "https://r4ds.had.co.nz/dates-and-times.html")
```

```{r, results = "asis", echo = FALSE}
extra_rec("")
```

* [A more in-depth discussion of the POSIXlt and POSIXct data classes.](https://www.cyclismo.org/tutorial/R/time.html)

* [A tutorial on lubridate](https://data.library.virginia.edu/working-with-dates-and-time-in-r-using-the-lubridate-package/) - scroll down for details on intervals if you have trouble with `%within%` and `%--%`

---

```{r, results = "asis", echo = FALSE}
checkin("Functions from lubridate")
```

* What is the difference between `year()` and `years()` and `dyears()`?

* What time zone does R think dates are in, by default?

* What is the difference between `%within%` and `%--%`?

* What happens if you use the `parse_datetime()` function, but you do not provide a `day` argument?

```{r, results = "asis", echo = FALSE}
checkin("Apply your skills")
```

**Question 1: Warm-up**

Answer the following using code (even if you know the answer already).

a. Save your birthday as a Datetime object in R.  

```{r, include = FALSE}
bday <- make_datetime(year = 1989, month = 04, day = 02)
```


b. What day of the week does/did your birthday fall on in 2020?

```{r, include = FALSE}
bday2 <- bday
year(bday2) <- 2020
bday2 %>% wday(label = TRUE)
```

c. On what day of the week should you plan your 100th birthday party?

```{r, include = FALSE}
(bday + years(100)) %>% wday(label = TRUE)
```

d. Find the date that is exactly 9 months before your birthday.  If it falls on a holiday, think about the implications.


**Question 2: Your Zodiac sign**

Download the `zodiac.csv` data from the course website.  This dataset gives the date ranges for the 12 Zodiac signs; that is, if you were born between the dates listed, you are affliated with that particular sign.

Use this dataset, and your new skills with DateTimes, to determine your sign *using only code*.
