---
title: "Animated Graphics in R"
subtitle: "Module 4"  
author: "Hunter Glanz & Kelly Bodwin"

output:
  xaringan::moon_reader:
    lib_dir: libs
    css: xaringan-themer.css
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
library(tidyverse)
library(palmerpenguins)
library(flair)
library(lubridate)
library(gganimate)
library(gifski)
```

```{r xaringan-themer, include=FALSE}
library(xaringanthemer)
solarized_light()
```

```{r, include = FALSE}
oscars <- read_csv("../../data/Oscars-demographics-DFE.csv")

oscars <- oscars %>%
  select(-matches("^_")) %>%
  select(-contains("gold")) %>%
  select(-contains("confidence")) %>%
  separate(date_of_birth,
           into = c("Day", "Month", "Year"),
           sep = "-") %>%
  mutate(
    Year = as.numeric(Year),
    Year = case_when(
      Year < 100 ~ Year + 1900,
      TRUE ~ Year
    ),
    Birthdate = ymd(paste(Year, Month, Day)),
    Date_of_award = ymd(paste(year_of_award, "Feb", "01")),
    Age_at_Award = interval(Birthdate, Date_of_award)/years(1),
    Award_Gender = case_when(
      str_detect(award, "Actress") ~ "Female",
      str_detect(award, "Actor") ~ "Male",
      TRUE ~ "Ungendered"
    ),
    Decade_of_Award = round(year_of_award, digits = -1)
  )

```

## gganimate

There are many, many ways to "spice up" your plots.  We will focus in on one: making your plot animated!

The best package for this, if you are using `ggplot` already, is `gganimate`.

*Note:* `gganimate` plot objects can sometimes take a long time to render.  One way to make it quicker is to change the number of frames in your gif.  Another trick is to use the [cache chunk option](https://bookdown.org/yihui/rmarkdown-cookbook/cache.html) in R Markdown, so that you don't re-render the images every time you knit your file.

---

## Our Dataset

A dataset concerning the race, religion, age, and other demographic details of all Oscar winners since 1928 the following categories:

* Best Actor
* Best Actress
* Best Supporting Actor
* Best Supporting Actress
* Best Director

---

## From Static to Animated

* The `transition_states()` function splits up plot data by a discrete variable and animates between the different states

```{r}
p <- ggplot(penguins, aes(x = bill_length_mm, y = bill_depth_mm)) +
  geom_point()

anim <- p +
  transition_states(species,
                    transition_length = 2,
                    state_length = 1)

anim
```

---

## Controlling the Movement

* *Easing* is the concept for how to display the intermediary data between the states specified in `transition_states()`.

* The default is linear, but there are other choices if you prefer

```{r}
anim +
  ease_aes("cubic-in-out") # slow start and end for a smoother look
```

---

## Annotating the Graphs

* Of course you will often want the axes, title, or other labels to change as the graph changes!

```{r}
anim +
  ggtitle('Now showing {closest_state}',
          subtitle = 'Frame {frame} of {nframes}')
```

---

## Object Permanence

* The penguin data animation here can be a bit confusing because we don't have the same number of penguins of each species. So watching one species morph into another isn't quite accurate.

* So instead of the data for each species "morphing" into the next, it may be more clear to eliminate the transition

```{r}
ggplot(penguins, aes(x = bill_length_mm, y = bill_depth_mm)) +
  geom_point(aes(colour = species)) +
  transition_states(species,
                    transition_length = 2,
                    state_length = 1)
```

---

## Sweet Sweet Style

* While the new transitions between species may be more accurate, we don't want to completely give up the spice and excitement!

* So let's add some back in...

```{r}
anim <- ggplot(penguins, aes(x = bill_length_mm, y = bill_depth_mm)) +
  geom_point(aes(colour = species), size = 2) +
  transition_states(species,
                    transition_length = 2,
                    state_length = 1)

anim +
  enter_fade() +
  exit_shrink()
```


---

## **Fill in the blanks**


```{r trans_state, include = FALSE, eval = FALSE}
p1 <- oscars %>%
  filter(Award_Gender != "Ungendered") %>%
  ggplot(aes(y = Age_at_Award, x = award, fill = Award_Gender)) +
  geom_boxplot() +
  ggtitle("Age of Academy Award Winners ({closest_state})") + 
  xlab("Year of Award") +
  ylab("Age at Award") +
  transition_states(Award_Gender) +
  shadow_mark(alpha = 0.3)

animate(p1, nframes = 10, fps = 5)
```

---

## The Result

```{r trans_state_res, include = FALSE}
p1 <- oscars %>%
  filter(Award_Gender != "Ungendered") %>%
  ggplot(aes(y = Age_at_Award, x = award, fill = Award_Gender)) +
  geom_boxplot() +
  ggtitle("Age of Academy Award Winners ({closest_state})") + 
  xlab("Year of Award") +
  ylab("Age at Award") +
  transition_states(Award_Gender) +
  shadow_mark(alpha = 0.3)

animate(p1, nframes = 10, fps = 5)
```

---

## **Fill in the blanks**

```{r trans_year, include = FALSE, eval = FALSE}
p2 <- oscars %>%
  filter(Award_Gender != "Ungendered") %>%
  ggplot(aes(x = Age_at_Award, fill = Award_Gender)) +
  geom_density(alpha = 0.5) +
  ggtitle("Age of Academy Award Winners ({frame_time}s)") + 
  xlab("Year of Award") +
  ylab("Age at Award") +
  transition_time(Decade_of_Award)

animate(p2, nframes = 9, fps = 1)
```

---

## The Result

```{r trans_year_res, include = FALSE}
p2 <- oscars %>%
  filter(Award_Gender != "Ungendered") %>%
  ggplot(aes(x = Age_at_Award, fill = Award_Gender)) +
  geom_density(alpha = 0.5) +
  ggtitle("Age of Academy Award Winners ({frame_time}s)") + 
  xlab("Year of Award") +
  ylab("Age at Award") +
  transition_time(Decade_of_Award)

animate(p2, nframes = 9, fps = 1)
```

---

## A Few More Options

```{r, echo = FALSE, cache = TRUE}
decorate("trans_state") %>%
  mask("closest_state", background = "#fdff32") %>%
  mask("states", background = "#fdff32") %>%
  mask("shadow_mark", background = "#fdff32") %>%
  mask("p1 ", background = "#fdff32") %>%
  mask("nframes", background = "#fdff32") %>%
  mask("fps", background = "#fdff32")


decorate("trans_year") %>%
  mask("p2 ", background = "#fdff32") %>%
  mask("transition_time", background = "#fdff32") %>%
  mask("frame_time", background = "#fdff32") %>%
  mask("nframes", background = "#fdff32") %>%
  mask("fps", background = "#fdff32")

```