---
title: "Exploration of Colleges Data"
output:
  html_document: default
  pdf_document: default
---
## Setup


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries}
library(tidyverse)
library(here)
```

```{r read_data}
colleges <- read_csv(here("data", "colleges_clean.csv"))
```


## Wrangling

```{r summarize}
#### Write code to show the mean and median and sd of in-state tution


```

```{r summarize2}
#### Write code to show the total undergrads in all institutions

```


```{r pipeline}
#### Write code to create the variable "tuition_diff" here


#### Write code to include only only relevant columns, and only colleges from the three states of interest


#### Write code to sort the dataset by tuition difference


#### Combine all steps into one pipeline
  
```

## Grouping and summarizing

* Find the average in-state tuition for public versus private schools.

* Find the average in-state tuition in NY, CA, and TX.


## Wrangling to Plot

Make the best plot you can to address the research question:

> Who has bigger in-state vs out-state difference: CA, Texas, or New York?

For bonus points, consider Public and Private school differences as well.

