---
title: "Exploration of Colleges Data"
output:
  html_document: default
  pdf_document: default
---
## Setup


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries}
library(tidyverse)
library(here)
```

```{r read_data}
colleges <- read_csv(here("data", "colleges_clean.csv"))
```


## Wrangling

```{r summarize}
#### Write code to show the mean and median and sd of in-state tuition

colleges %>%
  summarize(
    mean_tuit = mean(TUITIONFEE_IN),
    med_tuit = median(TUITIONFEE_IN),
    sd_tuit = sd(TUITIONFEE_IN)
  )

# OR

colleges %>%
  summarize_at(vars(TUITIONFEE_IN), 
               list(mean = mean, 
                    median = median, 
                    sd = sd))

```

```{r summarize2}
#### Write code to show the total undergrads in all institutions

colleges %>%
  summarize(sum(UGDS))

```


```{r pipeline}
#### Write code to create the variable "tuition_diff" here


#### Write code to include only relevant columns, and only colleges from the three states of interest


#### Write code to sort the dataset by tuition difference


#### Combine all steps into one pipeline

colleges_liminted <- colleges %>% 
  mutate(
    tuition_diff = TUITIONFEE_OUT - TUITIONFEE_IN
  ) %>%
  select(INSTNM, STABBR, CONTROL, tuition_diff) %>%
  filter(STABBR %in% c("NY", "CA", "TX")) %>%
  arrange(desc(tuition_diff))
  
```

## Grouping and summarizing

* Find the average in-state tuition for public versus private schools.

```{r}
colleges_limited %>%
  group_by(CONTROL) %>%
  summarize(TUITIONFEE_IN)
```


* Find the average in-state tuition in NY, CA, and TX.

```{r}
colleges_limited %>%
  group_by(STABBR) %>%
  summarize(TUITIONFEE_IN)
```



## Wrangling to Plot

Make the best plot you can to address the research question:

> Who has bigger in-state vs out-state difference: CA, Texas, or New York?

For bonus points, consider Public and Private school differences as well.

```{r}
## Left to you!
```


