---
title: "Introduction to Shiny"
subtitle: "Day 3"  
author: "Hunter Glanz & Kelly Bodwin"

output:
  xaringan::moon_reader:
    lib_dir: libs
    css: xaringan-themer.css
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
library(tidyverse)
```

```{r xaringan-themer, include=FALSE}
library(xaringanthemer)
solarized_light()
```

# What is Shiny?

* Shiny is a (package) framework for creating web applications using R code

--

* Perhaps more importantly, it's a relatively easy way to make your data work interactive and vastly more accessible!

--

* A nice preview on the [shiny gallery](https://shiny.rstudio.com/gallery/)

---

# A Bit Similar to R Projects

* Create a new folder on your desktop called `myfirstshinyapp`

* File -> New File -> Shiny Web App

  * Select the `app.R` option and specify your newly created folder `myfirstshinyapp`
  
  * This process creates a file called `app.R` (the names of these Shiny files are important; i.e. leave them as `app.R`, `ui.R`, and `server.R`)
  
  * Clear the entire contents of this new `app.R` file
  
  * Type the following into your new `app.R` file
  
```{r, eval = FALSE}
library(shiny)

ui <- "Hello, world!"

server <- function(input, output, session) {
}

shinyApp(ui, server)
```

---

# What Does This Do?

* Loads the `shiny` package

* Defines the user interface (UI) -- in this case just the words "Hello, world!"

* Specifies the behavior of our app by defining a `server` function. It's currently empty.

* Runs `shinyApp(ui, server)` to construct a Shiny application object from the UI object and server function

--

* Run this Shiny app by clicking the `Run App` button in the top right of the Script pane

---

# A Few Things of Note

* You can choose to view your running app in an RStudio Viewer pane/window or your system's default web browser (I prefer the browser)

* Notice what the R Console says right now...

--

  * R is busy! The R prompt isn't visible; you cannot run new code while the app is running
  
--

* Stop your Shiny app in one of the following ways:

1. Click the stop icon at the top of the R Console pane

2. Press the `Esc` key while the RStudio Console pane is selected

3. Close the Shiny app window

4. However you normally interrupt running R code, e.g. `Ctrl + C`

---

# Not Impressed?

* I'd be a little surprised if you were impressed!

--

* Our "Hello, world!" Shiny app isn't interactive and doesn't work with data or anything

* Y'all know what dataset I like?!

--

  * It's even got a sample [Shiny app](https://shiny.rstudio.com/gallery/kmeans-example.html)
  
* Let's talk through the basics of this `iris` example

---

# Why Reactivity

* Imagine we want to convert some temperatures:

```{r}
temp_c <- 10
temp_f <- (temp_c + 32) * 9 / 5 
temp_f
```

--

* But then, if we change `temp_c`

```{r}
temp_c <- 30
temp_f
```

---

# Functions Were a Previous Answer To This

```{r}
temp_c <- 10
temp_f <- function() {
  message("Converting") 
  (temp_c + 32) * 9 / 5 
}
temp_f()
```

* The value gets updated now, but we still have to call the `temp_f()` function EVERY time:

```{r}
temp_celsius <- -3
temp_f() 
```

---

# Reactive Programming Elegantly Solves This

1. We want up-to-date computations

2. We want to do as little computations as possible

* In other words...

--

* Expressions and outputs should update whenever one of their input values changes

* Expressions and outputs should *only* update when one of their input changes

--

* Shall we go back to the `iris` app to review?


---

# Overview of Inputs and Outputs

* Inputs

  * The [widget gallery](https://shiny.rstudio.com/gallery/widget-gallery.html)
  
--

* Outputs

  * All the plots you can think of (e.g. `ggplot` and more)
  
  * Interactive maps
  
  * Rich JavaScript output
  
  * Simple tables and text
  
  * Static pictures and gifs
  
  * And I'm sure much more!

---

# A Few More Examples

* [Classification Method Demonstration](https://mschroth.shinyapps.io/classificationapp/)

* [Astronaut Data Exploration](https://mschroth.shinyapps.io/lisinkershinyapp/)

* [General Data Exploration](https://ikamrath.shinyapps.io/DataExploration/)

---

class: inverse

# Shiny App Activity

* Create a new version of the `iris` example app that...

  * Takes the same two variable inputs (but not the number of clusters)
  
  * Outputs three things:
  
    * The same graph as before, but using ggplot (*Hint: use `aes_string` instead of `aes`*), colored by `Species`
    
    * A summary of the `x` variable
    
    * A summary of the `y` variable

